# Introduction  

## Background  

For newborns, current recommendations by the CDC are to primarily breastfeed a child for the first 6 months of their life and then move to gradually introduce solid foods into their diet. Previous studies have reported the mean and median breastfeeding times to be about 5.7 and 6 months respectively (Wasie Kasahun et al.). However, breastfeeding can continue for as long as the mother and child feel like they can continue for. In general, there are few drawbacks to breastfeeding for longer times, with some recommendations suggesting continuing breastfeeding for as long as 2 years (“Breastfeeding FAQs: How Much and How Often (for Parents) - Nemours KidsHealth”). Breastfeeding is considered beneficial for children, protecting them from certain diseases and past studies have shown that breastfed babies have lower risks of asthma, obesity, and sudden infant death syndrome (SIDS) (“Breastfeeding Benefits Both Baby and Mom | DNPAO | CDC”). While there are certainly multiple benefits to breastfeeding newborns, there's a lot of variability in the time until a mother chooses to wean her child. In some cases, mothers will only breastfeed their children for approximately a week before weaning them. The primary purpose of this analysis is to identify trends in the duration of the breastfeeding period as well as identify variables that appear to be associated with these trends.  

## Data Description  

The data was sourced from the `KMsurv` package in R. The original data is from the National Longitudinal Survey of Youth (a survey which began in 1979 and ended in 1988). The package data is a cleaned and subsetted version of the original data.  

Beginning in 1983, the women in the study population were asked about any pregnancies they had and various data about them were collected. For this dataset, only first-born children born after 1978 at a gestational age of 20 to 45 weeks were included. Then, the data was further narrowed down to only include responses from mothers who chose to breastfeed their children. After taking all of these factors into account, we end up with 927 total observations. For this analysis, the response (survival) is generated using the breastfeeding duration (weeks) and an indicator for whether or not the child was weaned.  

The variables included in this dataset are shown in Table \@ref(tab:facttab).   

```{r echo=FALSE, message=FALSE, warning=FALSE}
variable_names <- c("race", "poverty", "smoke", "alcohol", "agemth", "ybirth", "yschool", "pc3mth")
var_desc <- c("Race of mother (1=white, 2=black, 3=other)", 
"Mother in poverty (1=yes, 0=no)",
"Mother smoked at birth of child (1=yes, 0=no)",  
"Mother used alcohol at birth of child (1=yes, 0=no)",
"Age of mother at birth of child",
"Year of birth", 
"Education level of mother (years of school)",
"Prenatal care after 3rd month (1=yes, 0=no)")

variable_tab <- data.frame(variable_names, var_desc)
variable_tab %>% kable(col.names = NULL, 
                    align = 'c', 
                    booktabs = T, 
                    caption = "\\label{tab:facttab}Factor Levels for Categorical Variables") %>% 
  kable_styling(latex_options = c("striped", "HOLD_position"), 
                    full_width = F, font_size = 10 
                    ) %>% 
  column_spec(1, bold = T)
```

It is important to note that, for this dataset in particular, formula feeding is not considered breastfeeding despite formula being a form of liquid nourishment for infants. The indicator for poverty only indicates whether or not the mother was in poverty at the time of birth of her child. Table \@ref(tab:tab1) shows some very basic summary statistics for the outcome variables (`duration` and `delta`). When `delta` takes on a value of 1, it means that the child completed breastfeeding (was weaned) at the end of the reported period of time. Otherwise, the value of 0 meant that the information was either censored (unobserved) or there was loss to follow-up. All of the children who were reported to not have finished breastfeeding yet were born during or after 1984.   


```{r echo=FALSE, message=FALSE, warning=FALSE}
df %>% 
  tbl_summary(include = c(duration, delta), 
              type = all_continuous() ~ "continuous2",
              statistic = all_continuous() ~ c("{mean}", "{median}", "{min}-{max}"),
              label = list(duration = "duration (weeks)",
                           delta = "delta (=1)"))%>% 
  modify_header(label = "**bfeed Outcome Summary**") %>% 
  #modify_caption("\\label{tab:tab2}Summary of Response") %>% 
  as_kable_extra(format = "latex", booktabs = T,
                 caption = "\\label{tab:tab1}Summary of Data Outcome") %>% 
  kable_styling(latex_options = c("striped", "HOLD_position"), 
                    full_width = F
                    )
```

```{r echo=FALSE, message=TRUE, warning=FALSE}
df %>% 
  tbl_summary(include = c(race, yschool, poverty, alcohol, smoke, agemth, pc3mth), 
              type = all_continuous() ~ "continuous2",
              statistic = list(all_continuous() ~ c("{mean} ({sd})"),
                               all_categorical() ~ c("{n} ({p}%)")),
              by = delta) %>% 
  modify_header(label = "**bfeed Variable Summary**") %>% 
  #modify_caption("\\label{tab:tab3}Summary of Variables") %>% 
  as_kable_extra(format = "latex", booktabs = T,
                 caption = "\\label{tab:tab2}Summary of Data by Variable") %>% 
  kable_styling(latex_options = c("striped", "HOLD_position"), 
                    full_width = F
                    )
```

General guidelines for pregnant mothers recommend seeking prenatal care after the third month of pregnancy. However, as seen in Table \@ref(tab:tab2), only approximately 18% of the mothers chose to seek out prenatal care after the third month of their pregnancies. Also, while a fairly low percentage of the mothers consumed alcohol around the time of the birth of their child, approximately 30% of them reported smoking at the time. However, the degree of their smoking, such as how much or how often they smoked, wasn't included in the dataset. Visual representations of these summary tables can be found in the exploratory data analysis portion of the paper.   

